---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    mean_seats = mean(seats),
  ) |>
  arrange(desc(mean_seats))
```

**A1** Engineering, Cooperative Education has the highest mean number of seats. I'm surprised that this department has a lot more seats offered compared to the other departments. Personally, I hadn't realized that engineering was a popular subject.

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}
umd_courses |>
  group_by(id) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))
```

**A2** The greatest number of times a class has been offered is 22 times.

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}
umd_courses |>
  group_by(instructors, department) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(instructors, department)

```

**A3 Adam Nixon has the taught the most classes.**

------------------------------------------------------------------------

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

```{r}

```

**A4** It would be easier to see the list if there was a way to order it in descending order with the greatest number at the top.

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

```{r}
umd_courses |>
  group_by(term, id, department) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(term, id, department)
```

**A5** The data that I decided to test out was from my first question which was, **"**I would ask it to give me the number of courses for each department per semester. (Fall 2023, Spring 2024, etc.) I am curious to see the fluctuations in offerings for some classes and explore why a course isn't offered at a certain time." The data that I found was interesting. I focused on just one department, African American studies. Between 201712 and 201801, African American studies offered many more courses. The only courses that were offered for both terms were AASP100 and AASP202. Those two courses are probably introductory courses for students that are African American studies majors, minors or are just taking the classes as an elective. The other classes that were offered may just be classes for only African American studies majors or advanced courses for students minoring in this subject.
