---
title: "Reverse Engineering Project"
author: "Alaysia Ezzard"
date: "10/17/24"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data
fatal_police_shootings_agencies <- read.csv('fatal-police-shootings-agencies.csv') |> clean_names()

fatal_police_shootings_data <- read.csv('fatal-police-shootings-data.csv') |> clean_names()
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Of the 247 women fatally shot, 48 were Black and seven of those were unarmed.
-   **Analysis summary**: The result is almost the same. Our dataset was updated a few days ago, so the number of women fatally shot since 2015 has increased. Now there have been 358 women fatally shot, 58 of those women were Black and 7 were still unarmed.

```{r}
# Put code to reverse engineer sentence here
fatal_police_shootings_data |>
  filter(str_detect(armed,"unarmed")) |>
  group_by(armed, race, gender) |>
  summarize(total = n()) |>
    arrange(desc(total))

fatal_police_shootings_data |>
  filter(str_detect(race, "B")) |>
  group_by(gender,race) |>
  summarize(total=n()) |>
  arrange(desc(total)) 

fatal_police_shootings_data |>
  group_by(gender) |>
  summarize(total=n()) |>
  arrange(desc(total)) 

# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
